---
title: "Generating synthetic data using `stpp` package in R"
author: "geoMADE"
date: "19 October 2017"
output:
<<<<<<< HEAD
html_document: default
>>>>>>> c8704a4cfad98d706b7a23432d41bbf1fd63d1f1
---

NoTE: The github doesn't currently render the latex equations typed in 'rmarkdown' document. Therefore, in order to preview this document with equations properly rendered, follow the below steps:

1. Click on the 'DataSynth' directory
2. Click on the 'SyntheticDataInR.html' file, and copy the GitHub URL
3. Use this tool to preview it: http://htmlpreview.github.io/  (Credit: Jennifer Bryan)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

In R environment, the **stpp** package is the only package that provides the functions for synthesizing point process in space and time over a wide class of models. For the purpose of this post, we will look at three models, namely; the Homogeneous Poisson Process (HPP), Inhomogeneous Poisson Process (IHPP), and Poisson cluster process (PCP). These models are chosen because of their relevance from the viewpoint of crime data analytics. These will be discussed along with their formulations. 

The `rpp` function is used to synthesise point process based on the HPP and IHPP models, while the ` rpcp` function is used to synthesise point process based on the PCP model. In this exercise, the point processes are generated within a 3D region defined spatially by the geographical boundary of South Chicago, and temporally by one year time interval i.e. $t =c(1, 365)$. $$y_i = \alpha + \beta x_i + e_i$$.

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
